---
title: "Assignment 2, Task 2"
author: "Meghan Fletcher"
date: "2/5/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE, message = FALSE, warning = TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(broom)
library(janitor)
library(jtools)
library(kableExtra)
library(viridis)
```

## Part 1: Data Visualization
```{r}
# Read in the data
palmetto_data <- read_csv(here("data","palmetto.csv")) %>% 
  clean_names()

# Wrangle the data for visualization
palmetto_clean <- palmetto_data %>% 
  dplyr::select("species", "height", "length", "width", "green_lvs") %>% mutate(species = as.factor(species)) %>% 
  mutate(plant_name = case_when(
    species == 1 ~ "Serenoa repens",
    species == 2 ~ "Sabal etonia"
  )) %>% 
  mutate(plant_name = fct_drop(plant_name))
```

```{r, fig.cap = "**Figure 1:** The results represent the relationship between tree height and the amount of green leaves on both *Serenoa repens* and *Sabal etonia* (Data: Abrahamson et al. 2019)."}
# Create first visualization: height and green leaves comparison

ggplot(data = palmetto_clean, aes(x = height, y = green_lvs)) +
  geom_col(aes(fill=plant_name)) +
  facet_wrap(~ plant_name) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "Height (cm)", y = "Number of Green Leaves", title = "Leaf Content vs. Height of Serenoa repens and Sabal etonia", subtitle = "From 1982 to 2017") +
  scale_fill_manual(values = c("darkseagreen4", "chocolate4"))
```
#### From the graph above, we can see that both *Serenoa repens* and *Sabal etonia* appear to be similar in height (in centimaters). However, *Serenoa repens* tends to have a greater amount of green leaves. Green leaf amount appears to be greatest for trees with heights close to the mean and median values of height for both species. 


```{r, fig.cap="**Figure 2:** The points presented indicate that both species appear to have a strong, positive correlaiton between canopy length (cm) and canopy width (cm) (Data: Abrahamson et al. 2019)."}
# Create a second visualization looking at the relationship between canopy width and canopy length

ggplot(data = palmetto_clean, aes(x = length, y = width)) +
  geom_point(aes(color = plant_name), alpha = 0.7) +
  scale_color_manual(values = c("darkseagreen4", "chocolate4")) +
  geom_smooth(method='lm', color = "black") +
  facet_wrap(~plant_name) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "Canopy Length (cm)", y = "Canopy Width (cm)", title = "Compariosn of Canopy Length vs. Canopy Width of Serenoa repens and Sabal etonia", subtitle = "From 1982 to 2017") 
  
    
```

#### As canopy length (cm) increases, canopy width (cm) appears to increase almost evenly for both *Serenoa repens* and *Sabal etonia*.

```{r, fig.cap="**Figure 3:** This boxplot shows us that based on the average canopy length (cm) of both species, *Serenoa repens* typically has a greater amount of green leaves present as compared to *Sabal etonia* (Data: Abrahamson et al. 2019)."}

# Create third visualization looking at amount of green leaves by species

ggplot(data = palmetto_clean, aes(x = length, y = green_lvs)) +
  geom_boxplot(aes(fill = plant_name)) +
  facet_wrap(~plant_name) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "Canopy Length (cm)", y = "Amount of Green Leaves", title = "Green Leaf Content of Serenoa repens and Sabal etonia", subtitle = "From 1982 to 2017") +
  scale_fill_manual(values = c("darkseagreen4", "chocolate4"))
```

#### Based on canopy length (cm), *Serenoa repens* appears to have a greater amount of green leaves on average than *Sabal etonia*. This result may change if the amount of green leaves are based on height or canopy width.



## Part 2: Binary Logistic Regression

#### Priliminary binary logitic regression results:
```{r}
# create a binary logistic regression model and table
palmetto_blr <- glm(plant_name ~ height + length + width + green_lvs,
                    data = palmetto_clean,
                    family = "binomial")

summary(palmetto_blr)
```

##### **Table 1:** The data presented depicts the preliminary results from a binary logitic regression model performed on the Palmetto data where *Serenoa repens* and *Sabal etonia* are the plant species and the four predictor variables used are plant height (cm), canopy length (cm), canopy width (cm) and amount of green leaves present (Data: Abrahamson et al. (2019)).
```{r}
# Tidy up the data and create a table
palmetto_blr_tidy <- tidy(palmetto_blr)

palmetto_blr_tidy %>% 
  kbl(caption = "Palmetto Binary Logistic Regression Model Results") %>% 
   kable_classic_2(full_width = F, html_font = "Cambria")
```

## Part 3: Finding the Probability of Predicting the Correct Plant Species using BLR
```{r}
palmetto_blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response")
```

##### **Table 2:** The results display the correctness of the binary logistic regression model used. The percentages of *Serenoa repens* and *Sabal etonia* that were successfully and unsuccessfully predicted by the model based on the four predictor variables are displayed as well, including the actual number of each outcome in parantheses (Data: Abrahamson et al. (2019)). 
```{r}
# Create a fitted table
palmetto_fitted_table <- palmetto_blr_fitted %>% 
  mutate(Predicted = case_when(
    .fitted > .50 ~ "Serenoa repens",
    .fitted <= .50 ~ "Sabal etonia"
  )) 

palmetto_final <- palmetto_fitted_table %>% 
  mutate(Correct = case_when(Predicted == plant_name ~ "Plants Successfully Predicted",
                             TRUE ~ "Plants Unsuccessfully Predicted")) %>% 
  select(plant_name, .fitted, Predicted, Correct) %>% 
  group_by(plant_name, Correct) %>% 
  summarize(Results = n()) %>% 
  pivot_wider(plant_name, names_from = "Correct",
              values_from = "Results") %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns() %>% 
  rename("Plant Species" = plant_name)

palmetto_final %>% 
  kbl(caption = "Palmetto Plant Species BInary Logitic Regression Prediction Outcomes") %>% 
   kable_classic_2(full_width = F, html_font = "Cambria")
```

#### Overall, the BLR model used did a decent job at predicting the correct plant speices the majority of the time. 


#### **Data citation:** Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5
